High order finite volume operators in a cut cell context, Part 1: system matrix conditioning for finite volume elliptic operators

### 2023-07-07 example/EBCM/hoeb_petsc/_doc/paper.tex
added latest 3d data
millions of typos cry out at once

\title{Accuracy and stability for elliptic operators
  in a finite volume context}
\author{
   \and D. Devendran    \footnotemark[2]
   \and D. T. Graves    \footnotemark[1]
        }

@@

Devendran, et al. \cite{Devendran2017} present an algorithm that
solves Poisson's equation to fourth-order accuracy using an EB grid.
Generalizing somewhat  the framework they present, we can define a
family of elliptic operators.
Here we investigate how the accuracy and stability of this family
of operators varies within its parameter space.

@@

Embedded boundary (EB) grids are formed when one passes an  surface
through a Cartesian mesh.    For sufficiently complex geomtries, these methods
are very attactive because grid generation is a solved problem
even for moving grids \cite{MillerTrebotich2012}.
In the current context, we form the cutting surface as the zero
surface of a function  of space $I(\xbold)$, $\xbold \in R^D$.
For smooth ($I$),  moments can be generated to any
accuracy \cite{Schwartz2015}.

@@

We approximate $\psi$
in the neighborhood of $\xbar$ using a Taylor expansion to order $P_T$:
\begin{equation}
\psi(\xbold)  =  \sum\limits_{p < P_T} C^p (\xbold -\xbar)^p
\label{eqn::taylor}
\end{equation}
where $C^p$ this appropriate Taylor coefficient.  In three dimensions,
\begin{equation}
  C^p =\frac{1}{p!}
      \frac{\partial^{p_0}}{\partial x_0}
      \frac{\partial^{p_1}}{\partial x_1}
      \frac{\partial^{p_2}}{\partial x_2}  (\psi).
\end{equation}
If we know the local Taylor coeffiencts,
the gradient of the field  in direction $d$ is given by
\begin{equation}
  \frac{\partial \psi}{\partial x^d}  =
  \sum\limits_{p < P_T} p^d C^{p} (\xbar -\xbar)^{p-e^d}
\end{equation}

@@

For each $\cbold$ within the $\ncal_\fbold$, we get an equation of the form
\begin{equation}
  \int\limits_{A(\cbold)}  \nabla \psi \cdot \nhat dA =
  \sum\limits_{d < D} \int\limits_{A(\cbold)} \sum_{p < P} p_d C^p
  \xminxbar^{p-e^d} dA
  = 
\sum\limits_{d < D} \sum_{p < P}  p_d C_p \ecal^{p-e^d,d}_\cbold = 0.
\end{equation}
We force there to be enough neighbors so that this system is
overdetermined.  Let's say $\ncal_\fbold$ contains $N_B$ boundary faces and
$N_V$ volumes.   To solve for Taylor coefficients, we have a system 
of the form
\begin{equation*}
M C = P 
\end{equation*}
where $C=\{C^p\}$,

\begin{equation}
  M_\fbold =\left\{
    \begin{array}{c} 
       m^p_\jbold  \\
       p^d b^{p-e^d}_\bbold  \\
       p^d c^{p-e^d}_\cbold  
    \end{array}
 \right\},
\end{equation}
and
\begin{equation}
  P_\fbold =
    \left\{
    \begin{array}{c} 
       <\psi>_\jbold  \\
      0  \\
      0  
    \end{array}
    \right\}
\end{equation}
for every volume $\jbold$, cut face $\cbold$ and domain face $\bbold$
in $\ncal_\fbold$.
Since the system is overdetermined, we can introduce a 
meaningful weighting matrix $W$ \footnotemark[4]. 
 Choice of $W$
is crucial to the stability of these algorithms
\footnotemark[5].   
We multiply and solve
\begin{equation*}
WMC = WP.
\end{equation*}
Since the system is (deliberately) not square, we approximate
the solution using a Moore-Penrose pseudo-inverse:
\begin{equation}
  C = ((WM)^T(WM))^{-1} (WM)^T P.
\label{eqn::getC}  
\end{equation}
where $A \equiv ((WM)^T(WM))_\fbold$ is the system matrix for face
$\fbold$.

@@

All volumes and faces  start with an  index $\ibold \in Z^D$.
We start from face $\fbold$.   Say we want to get
the distance from $\fbold$ with index $\ibold$ to a volume $V_1$
with index $\vbold$.  The distance $D_\fbold(V_1)$ is given by
the Manhattan disance:
\begin{equation}
  D_\fbold(V_1) = \sum\limits_{0 <= d < D} |\vbold^d - \ibold_0^d|.
\label{eqn::uptownFunk}
\end{equation}
$\gbold_\fbold(V_1)$, the weight of the $V_1$ equation in $\fbold$ system
is given by
\begin{equation}
  \gbold_\fbold(V_1)  \equiv \left(\frac{1}{D_\fbold(V_1) } \right)^{P_w},
\label{eqn::weightFunk}  
\end{equation}
where we call $P_w$ the weighting exponent.
The weighting matrix $W_\fbold = \{\{\wbold_{i,j}\}\}$ is diagonal
\begin{equation}
w_{i,j} =
\left(
\begin{array}{ll}
 \gbold_\fbold(V_i) & \mbox{ if } $(i==j)$ \\
 0   & \mbox{ otherwise }
\end{array}
\right)
\end{equation}
Once one has the Taylor coefficients, she can compute the flux through
the open face using  equation \ref{eqn::faceavg}.   Put into matrix
form, we define  vector  $Q_\fbold = \{q^p_\fbold\}$:
\begin{equation}
q^p_\fbold = p^d \mcal^{p-e^d}_\fbold
\end{equation}
and the flux at the face is given by
\begin{equation}
<\nabla \psi>_\fbold = Q C P= Q ((WM)^T(WM))^{-1} (WM)^T  P.
\end{equation}

@@

Overton-Katz, et al. \cite{Katz2023} present the above procedure in
the absence of boundary condition equations.   They show that, for
certain values of $P_T$ and $P_N$, the resulting $A$ matrix can have a
severe mismatch between its largest and smallest eigenvalues,
signalling that $A$ is poorly conditioned.    Their data show the
worst solvability issues occur near the domain boundary.    It stands to
reason that including boundary condition equations could improve the
conditioning of the resultant matricies. To test this, we present the
same measurement of condition number for  several values of $P_T$ and
$P_N$ where boundary condition equations are included when they are
part of the neighborhood.    We present the results for the same three
different geometries used in \cite{Katz2023}.  In figure \ref{fig::allRegSolvability} we show
the worst condition numbers where there are no cut cells.
In figure \ref{fig::unmergedSolvability}, we show the worst condition
number where we cut the unit domain with a centered sphere of radius
0.45 (this geometry includes very small volumes).    Finally, in
figure \ref{fig::mergedSolvability}, we present results all small
volumes have been merged with neighbors.

The data paint a very consistent picture,   Though adding boundary
condition equations does improve the solvability of $A$, aggressive
choices of $P_W$ and $P_T$ can lead to system matricies that cannot be
credibly solved without very high precision numerical arithmetic.
\begin{small}
\begin{table}
\begin{center}
\begin{tabular}{|ccc|ccc|} \hline
 $D$ & $P^T$  & $P^W$  & $\lambda_{max}$ & $\lambda_{min}$   & $\cali$ \\
 \hline
2 & 1 & 1 &  1.537125e+00 & 1.150772e-03 & 7.486517e-04 \\
2 & 1 & 2 &  4.940953e-01 & 9.196245e-05 & 1.861229e-04 \\
2 & 1 & 3 &  3.324177e-01 & 1.064946e-05 & 3.203639e-05 \\
2 & 1 & 4 &  2.853322e-01 & 1.705230e-06 & 5.976295e-06 \\
2 & 1 & 5 &  2.665812e-01 & 3.349335e-07 & 1.256403e-06 \\
2 & 2 & 1 &  2.407807e+00 & 1.457063e-06 & 6.051408e-07 \\
2 & 2 & 2 &  4.940956e-01 & 5.469314e-08 & 1.106934e-07 \\
2 & 2 & 3 &  3.324177e-01 & 2.879583e-09 & 8.662545e-09 \\
2 & 2 & 4 &  2.853323e-01 & 1.628939e-10 & 5.708921e-10 \\
2 & 2 & 5 &  2.665812e-01 & 1.084310e-11 & 4.067466e-11 \\
2 & 3 & 1 &  2.407807e+00 & 5.822902e-10 & 2.418342e-10 \\
2 & 3 & 2 &  9.272889e-01 & 4.520282e-11 & 4.874729e-11 \\ 
2 & 3 & 3 &  3.324177e-01 & 1.618058e-12 & 4.867544e-12 \\
2 & 3 & 4 &  2.853323e-01 & 7.227717e-14 & 2.533088e-13 \\
2 & 3 & 5 &  2.665812e-01 & 3.194386e-15 & 1.198279e-14 \\
2 & 4 & 1 &  1.880177e+00 & 1.535955e-25 & 8.169205e-26 \\
2 & 4 & 2 &  9.272889e-01 & 1.098665e-27 & 1.184814e-27 \\
2 & 4 & 3 &  6.561304e-01 & 5.104096e-28 & 7.779089e-28 \\
2 & 4 & 4 &  2.868930e-01 & 3.731928e-27 & 1.300808e-26 \\
2 & 4 & 5 &  5.327046e-01 & 5.118416e-25 & 9.608357e-25 \\
\hline
3 & 1 & 1 &  4.639099e+00 & 3.887142e-03 & 8.379088e-04 \\ 
3 & 1 & 2 &  1.079504e+00 & 2.369894e-04 & 2.195354e-04 \\ 
3 & 1 & 3 &  6.747118e-01 & 2.359590e-05 & 3.497182e-05 \\ 
3 & 1 & 4 &  5.721834e-01 & 3.540976e-06 & 6.188534e-06 \\ 
3 & 1 & 5 &  5.333668e-01 & 6.792558e-07 & 1.273525e-06 \\ 
3 & 2 & 1 &  4.639156e+00 & 3.176883e-06 & 6.847975e-07 \\ 
3 & 2 & 2 &  1.079505e+00 & 1.512048e-07 & 1.400686e-07 \\ 
3 & 2 & 3 &  6.747119e-01 & 6.991076e-09 & 1.036157e-08 \\ 
3 & 2 & 4 &  5.721835e-01 & 3.641527e-10 & 6.364265e-10 \\ 
3 & 2 & 5 &  2.669165e-01 & 1.146071e-11 & 4.293744e-11 \\ 
3 & 3 & 1 &  4.639157e+00 & 1.340233e-09 & 2.888959e-10 \\ 
3 & 3 & 2 &  1.079505e+00 & 7.151353e-11 & 6.624659e-11 \\ 
3 & 3 & 3 &  6.747119e-01 & 3.977531e-12 & 5.895155e-12 \\ 
3 & 3 & 4 &  2.869943e-01 & 5.757645e-14 & 2.006188e-13 \\ 
3 & 3 & 5 &  2.669165e-01 & 1.723124e-15 & 6.455665e-15 \\ 
3 & 4 & 1 &  6.402997e+00 & 5.729285e-26 & 8.947817e-27 \\ 
3 & 4 & 2 &  1.079505e+00 & 1.226846e-27 & 1.136489e-27 \\ 
3 & 4 & 3 &  6.907344e-01 & 1.049444e-28 & 1.519317e-28 \\ 
3 & 4 & 4 &  2.903904e-01 & 8.442770e-26 & 2.907386e-25 \\ 
3 & 4 & 5 &  5.333668e-01 & 5.987514e-25 & 1.122588e-24 \\ 
\hline
\end{tabular}
\end{center}
\label{fig::allRegSolvability}
\caption
    {
      We show the lowest value of
      $\cali$ for an entire Cartesian grid of size $32^D$
      with all
      boundary conditions homogeneous Neumann.
      We vary Taylor power $P^T$, weighting
      function exponents $P^W$, and dimensionality $D$. 
    }
\end{table}
\end{small}

@@

For context, allow us to recap a relevant bit of finite volume formalism.
Consider a volume $\ibold$ bounded by the set of faces
$\{f\}  = \partial \ibold$, where the outward-facing normal $\nhat$ of
each face is known, as is its  $V_\ibold$.
Given a flux function $F$, the average divergence $< \nabla \cdot F >$
is given by the divergence theorem:
\begin{equation}
\begin{array}{l l}
< \nabla \cdot F > &= \frac{1}{V_\ibold} \int_\ibold \nabla \cdot F dV \\
           &= \frac{1}{V_\ibold} \int_{\partial \ibold} F\cdot \nhat dA \\
           &= \frac{1}{V_\ibold} \sum_{f \in \partial
  \ibold}\int_{\partial \ibold} F\cdot \nhat dA.
\end{array}
\label{eqn::discreteDivergence}
\end{equation}
For all coordinate-alligned faces, the section above's description for how to
compute the flux is adequate.   What remains are fluxes through cut
faces.

@@

\subsection{The effect of Dirichlet boundary conditions on
  system matrix conditioning} In section \ref{sec::openface}, we
discuss how adding Neumann boundary condition equations can improve
the conditioning of the system matrix.  When the boundary conditions
are set to Dirichlet, we get similar results.  See tables
\ref{fig::mergedDiri}, \ref{fig::unmergedDiri}, and
\ref{fig::allRegDiri} for details.  Again, boundary conditions help
stabilize the system but it is still the case that agressive choices
of $P_T$ or $P_W$ can result in system matrices that cannot be
credibly inverted without very high precision numerical arithmetic.
Again, this seems less a matter of a particular grid configuration and
more  a property of the mathematical framework used to generate the stencil.
\begin{small}
\begin{table}
\begin{center}
\begin{tabular}{|ccc|ccc|} \hline
 $D$ & $P^T$  & $P^W$  & $\lambda_{max}$ & $\lambda_{min}$   & $\cali$ \\
 \hline
2 & 1 & 1 &  1.537125e+00 & 1.150772e-03 & 7.486517e-04 \\ 
2 & 1 & 2 &  4.940953e-01 & 9.196245e-05 & 1.861229e-04 \\ 
2 & 1 & 3 &  3.324177e-01 & 1.064946e-05 & 3.203639e-05 \\ 
2 & 1 & 4 &  2.853322e-01 & 1.705230e-06 & 5.976295e-06 \\ 
2 & 1 & 5 &  2.665812e-01 & 3.349335e-07 & 1.256403e-06 \\ 
2 & 2 & 1 &  2.407807e+00 & 1.457063e-06 & 6.051408e-07 \\ 
2 & 2 & 2 &  4.940956e-01 & 5.469314e-08 & 1.106934e-07 \\ 
2 & 2 & 3 &  3.324177e-01 & 2.879583e-09 & 8.662545e-09 \\ 
2 & 2 & 4 &  2.853323e-01 & 1.628939e-10 & 5.708921e-10 \\ 
2 & 2 & 5 &  2.665812e-01 & 1.084310e-11 & 4.067466e-11 \\ 
2 & 3 & 1 &  2.407807e+00 & 5.822902e-10 & 2.418342e-10 \\ 
2 & 3 & 2 &  9.272889e-01 & 4.520282e-11 & 4.874729e-11 \\ 
2 & 3 & 3 &  3.324177e-01 & 1.618058e-12 & 4.867544e-12 \\ 
2 & 3 & 4 &  2.853323e-01 & 7.227717e-14 & 2.533088e-13 \\ 
2 & 3 & 5 &  2.665812e-01 & 3.194386e-15 & 1.198279e-14 \\ 
2 & 4 & 1 &  2.943604e+00 & 6.445948e-26 & 2.189815e-26 \\ 
2 & 4 & 2 &  9.877639e-01 & 6.302117e-27 & 6.380185e-27 \\ 
2 & 4 & 3 &  6.561304e-01 & 6.778809e-28 & 1.033150e-27 \\ 
2 & 4 & 4 &  2.868930e-01 & 5.733222e-27 & 1.998383e-26 \\ 
2 & 4 & 5 &  5.327046e-01 & 5.117605e-25 & 9.606835e-25 \\
\hline
3 & 1 & 1 &  4.639099e+00 & 3.887142e-03 & 8.379088e-04 \\ 
3 & 1 & 2 &  1.079504e+00 & 2.369894e-04 & 2.195354e-04 \\ 
3 & 1 & 3 &  6.747118e-01 & 2.359590e-05 & 3.497182e-05 \\ 
3 & 1 & 4 &  5.721834e-01 & 3.540976e-06 & 6.188534e-06 \\ 
3 & 1 & 5 &  5.333668e-01 & 6.792558e-07 & 1.273525e-06 \\ 
3 & 2 & 1 &  4.639156e+00 & 3.176883e-06 & 6.847975e-07 \\ 
3 & 2 & 2 &  1.079505e+00 & 1.512048e-07 & 1.400686e-07 \\ 
3 & 2 & 3 &  6.747119e-01 & 6.991076e-09 & 1.036157e-08 \\ 
3 & 2 & 4 &  5.721835e-01 & 3.641527e-10 & 6.364265e-10 \\ 
3 & 2 & 5 &  2.669165e-01 & 1.146071e-11 & 4.293744e-11 \\ 
3 & 3 & 1 &  4.639157e+00 & 1.340233e-09 & 2.888959e-10 \\ 
3 & 3 & 2 &  1.079505e+00 & 7.151353e-11 & 6.624659e-11 \\ 
3 & 3 & 3 &  6.747119e-01 & 3.977531e-12 & 5.895155e-12 \\ 
3 & 3 & 4 &  2.869943e-01 & 5.757645e-14 & 2.006188e-13 \\ 
\hline
\end{tabular}
\end{center}
\label{fig::allRegDiri}
\caption
    {
      We show the lowest value of
      $\cali$ for an entire Cartesian grid of size $32^D$ with all
      boundary conditions homogeneous Dirichlet.
      We vary Taylor power $P^T$, weighting
      function exponents $P^W$, and dimensionality $D$.
%%(as of 6-28-2023, some 3d results still pending).      
    }
\end{table}
\end{small}

>>>>

\title{High order finite volume operators in a cut cell context,
  Part 1: system  matrix conditioning for finite volume elliptic operators}
\author{
   \and D. Devendran    \footnotemark[2]
   \and D. T. Graves    \footnotemark[1]
        }

@@

Devendran, et al. \cite{Devendran2017} present an algorithm that
solves Poisson's equation to fourth-order accuracy using an EB grid.
Generalizing somewhat  the framework they present, we can define a
family of elliptic operators.   That algorithm is a special case of a
family of finite volume elliptic operators.   We present the formal
process to generate these operators which includes inverting a system
matrix.    Overton-Katz, et al. \cite{Katz2023}, show that the system
matrix can have solvability problems in the absence of boundary
condition equations in the system matrix  but Devendran uses
boundary condition equations.   We present data that shows these extra
equations can improve the system matrix condition number.
We also show that, for certain parameters of operator,
this system matrix is still quite poorly conditioned.

@@

Embedded boundary (EB) grids are formed when one passes an  surface
through a Cartesian mesh.    For sufficiently complex geometries, these methods
are very attractive because grid generation is a solved problem
even for moving grids \cite{MillerTrebotich2012}.
In the current context, we form the cutting surface as the zero
surface of a function  of space $I(\xbold)$, $\xbold \in R^D$.
For smooth ($I$),  moments can be generated to any
accuracy \cite{Schwartz2015}.

@@

We approximate $\psi$
in the neighborhood of $\xbar$ using a Taylor expansion to order $P_T$:
\begin{equation}
\psi(\xbold)  =  \sum\limits_{p < P_T} C^p (\xbold -\xbar)^p
\label{eqn::taylor}
\end{equation}
where $C^p$ this appropriate Taylor coefficient.  In three dimensions,
\begin{equation}
  C^p =\frac{1}{p!}
      \frac{\partial^{p_0}}{\partial x_0}
      \frac{\partial^{p_1}}{\partial x_1}
      \frac{\partial^{p_2}}{\partial x_2}  (\psi).
\end{equation}
If we know the local Taylor coefficients,
the gradient of the field  in direction $d$ is given by
\begin{equation}
  \frac{\partial \psi}{\partial x^d}  =
  \sum\limits_{p < P_T} p^d C^{p} (\xbar -\xbar)^{p-e^d}
\end{equation}

@@

For each $\cbold$ within the $\ncal_\fbold$, we get an equation of the form
\begin{equation}
  \int\limits_{A(\cbold)}  \nabla \psi \cdot \nhat dA =
  \sum\limits_{d < D} \int\limits_{A(\cbold)} \sum_{p < P} p_d C^p
  \xminxbar^{p-e^d} dA
  = 
\sum\limits_{d < D} \sum_{p < P}  p_d C_p \ecal^{p-e^d,d}_\cbold = 0.
\end{equation}
We force there to be enough neighbors so that this system is
over-determined.  Let's say $\ncal_\fbold$ contains $N_B$ boundary faces and
$N_V$ volumes.   To solve for Taylor coefficients, we have a system 
of the form
\begin{equation*}
M C = P 
\end{equation*}
where $C=\{C^p\}$,

\begin{equation}
  M_\fbold =\left\{
    \begin{array}{c} 
       m^p_\jbold  \\
       p^d b^{p-e^d}_\bbold  \\
       p^d c^{p-e^d}_\cbold  
    \end{array}
 \right\},
\end{equation}
and
\begin{equation}
  P_\fbold =
    \left\{
    \begin{array}{c} 
       <\psi>_\jbold  \\
      0  \\
      0  
    \end{array}
    \right\}
\end{equation}
for every volume $\jbold$, cut face $\cbold$ and domain face $\bbold$
in $\ncal_\fbold$.
Since the system is over-determined, we can introduce a 
meaningful weighting matrix $W$ \footnotemark[4]. 
 Choice of $W$
is crucial to the stability of these algorithms
\footnotemark[5].   
We multiply and solve
\begin{equation*}
WMC = WP.
\end{equation*}
Since the system is (deliberately) not square, we approximate
the solution using a Moore-Penrose pseudo-inverse:
\begin{equation}
  C = ((WM)^T(WM))^{-1} (WM)^T P.
\label{eqn::getC}  
\end{equation}
where $A \equiv ((WM)^T(WM))_\fbold$ is the system matrix for face
$\fbold$.

@@

All volumes and faces  start with an  index $\ibold \in Z^D$.
We start from face $\fbold$.   Say we want to get
the distance from $\fbold$ with index $\ibold$ to a volume $V_1$
with index $\vbold$.  The distance $D_\fbold(V_1)$ is given by
the Manhattan distance:
\begin{equation}
  D_\fbold(V_1) = \sum\limits_{0 <= d < D} |\vbold^d - \ibold_0^d|.
\label{eqn::uptownFunk}
\end{equation}
$\gbold_\fbold(V_1)$, the weight of the $V_1$ equation in $\fbold$ system
is given by
\begin{equation}
  \gbold_\fbold(V_1)  \equiv \left(\frac{1}{D_\fbold(V_1) } \right)^{P_w},
\label{eqn::weightFunk}  
\end{equation}
where we call $P_w$ the weighting exponent.
The weighting matrix $W_\fbold = \{\{\wbold_{i,j}\}\}$ is diagonal
\begin{equation}
w_{i,j} =
\left(
\begin{array}{ll}
 \gbold_\fbold(V_i) & \mbox{ if } $(i==j)$ \\
 0   & \mbox{ otherwise }
\end{array}
\right)
\end{equation}
Once one has the Taylor coefficients, she can compute the flux through
the open face using  equation \ref{eqn::faceavg}.   Put into matrix
form, we define  vector  $Q_\fbold = \{q^p_\fbold\}$:
\begin{equation}
q^p_\fbold = p^d \mcal^{p-e^d}_\fbold
\end{equation}
and the flux at the face is given by
\begin{equation}
<\nabla \psi>_\fbold = Q C P= Q ((WM)^T(WM))^{-1} (WM)^T  P.
\end{equation}

@@

Overton-Katz, et al. \cite{Katz2023} present the above procedure in
the absence of boundary condition equations.   They show that, for
certain values of $P_T$ and $P_N$, the resulting $A$ matrix can have a
severe mismatch between its largest and smallest eigenvalues,
signaling that $A$ is poorly conditioned.    Their data show the
worst solvability issues occur near the domain boundary.    It stands to
reason that including boundary condition equations could improve the
conditioning of the resultant matrices. To test this, we present the
same measurement of condition number for  several values of $P_T$ and
$P_N$ where boundary condition equations are included when they are
part of the neighborhood.    We present the results for the same three
different geometries used in \cite{Katz2023}.  In figure \ref{fig::allRegSolvability} we show
the worst condition numbers where there are no cut cells.
In figure \ref{fig::unmergedSolvability}, we show the worst condition
number where we cut the unit domain with a centered sphere of radius
0.45 (this geometry includes very small volumes).    Finally, in
figure \ref{fig::mergedSolvability}, we present results all small
volumes have been merged with neighbors.

The data paint a very consistent picture,   Though adding boundary
condition equations does improve the solvability of $A$, aggressive
choices of $P_W$ and $P_T$ can lead to system matrices that cannot be
credibly solved without very high precision numerical arithmetic.
\begin{small}
\begin{table}
\begin{center}
\begin{tabular}{|ccc|ccc|} \hline
 $D$ & $P^T$  & $P^W$  & $\lambda_{max}$ & $\lambda_{min}$   & $\cali$ \\
 \hline
2 & 1 & 1 &  1.537125e+00 & 1.150772e-03 & 7.486517e-04 \\
2 & 1 & 2 &  4.940953e-01 & 9.196245e-05 & 1.861229e-04 \\
2 & 1 & 3 &  3.324177e-01 & 1.064946e-05 & 3.203639e-05 \\
2 & 1 & 4 &  2.853322e-01 & 1.705230e-06 & 5.976295e-06 \\
2 & 1 & 5 &  2.665812e-01 & 3.349335e-07 & 1.256403e-06 \\
2 & 2 & 1 &  2.407807e+00 & 1.457063e-06 & 6.051408e-07 \\
2 & 2 & 2 &  4.940956e-01 & 5.469314e-08 & 1.106934e-07 \\
2 & 2 & 3 &  3.324177e-01 & 2.879583e-09 & 8.662545e-09 \\
2 & 2 & 4 &  2.853323e-01 & 1.628939e-10 & 5.708921e-10 \\
2 & 2 & 5 &  2.665812e-01 & 1.084310e-11 & 4.067466e-11 \\
2 & 3 & 1 &  2.407807e+00 & 5.822902e-10 & 2.418342e-10 \\
2 & 3 & 2 &  9.272889e-01 & 4.520282e-11 & 4.874729e-11 \\ 
2 & 3 & 3 &  3.324177e-01 & 1.618058e-12 & 4.867544e-12 \\
2 & 3 & 4 &  2.853323e-01 & 7.227717e-14 & 2.533088e-13 \\
2 & 3 & 5 &  2.665812e-01 & 3.194386e-15 & 1.198279e-14 \\
2 & 4 & 1 &  1.880177e+00 & 1.535955e-25 & 8.169205e-26 \\
2 & 4 & 2 &  9.272889e-01 & 1.098665e-27 & 1.184814e-27 \\
2 & 4 & 3 &  6.561304e-01 & 5.104096e-28 & 7.779089e-28 \\
2 & 4 & 4 &  2.868930e-01 & 3.731928e-27 & 1.300808e-26 \\
2 & 4 & 5 &  5.327046e-01 & 5.118416e-25 & 9.608357e-25 \\
\hline
3 & 1 & 1 &  4.639099e+00 & 3.887142e-03 & 8.379088e-04 \\ 
3 & 1 & 2 &  1.079504e+00 & 2.369894e-04 & 2.195354e-04 \\ 
3 & 1 & 3 &  6.747118e-01 & 2.359590e-05 & 3.497182e-05 \\ 
3 & 1 & 4 &  5.721834e-01 & 3.540976e-06 & 6.188534e-06 \\ 
3 & 1 & 5 &  5.333668e-01 & 6.792558e-07 & 1.273525e-06 \\ 
3 & 2 & 1 &  4.639156e+00 & 3.176883e-06 & 6.847975e-07 \\ 
3 & 2 & 2 &  1.079505e+00 & 1.512048e-07 & 1.400686e-07 \\ 
3 & 2 & 3 &  6.747119e-01 & 6.991076e-09 & 1.036157e-08 \\ 
3 & 2 & 4 &  5.721835e-01 & 3.641527e-10 & 6.364265e-10 \\ 
3 & 2 & 5 &  2.669165e-01 & 1.146071e-11 & 4.293744e-11 \\ 
3 & 3 & 1 &  4.639157e+00 & 1.340233e-09 & 2.888959e-10 \\ 
3 & 3 & 2 &  1.079505e+00 & 7.151353e-11 & 6.624659e-11 \\ 
3 & 3 & 3 &  6.747119e-01 & 3.977531e-12 & 5.895155e-12 \\ 
3 & 3 & 4 &  2.869943e-01 & 5.757645e-14 & 2.006188e-13 \\ 
3 & 3 & 5 &  2.669165e-01 & 1.723124e-15 & 6.455665e-15 \\ 
3 & 4 & 1 &  6.402997e+00 & 5.729285e-26 & 8.947817e-27 \\ 
3 & 4 & 2 &  1.079505e+00 & 1.226846e-27 & 1.136489e-27 \\ 
3 & 4 & 3 &  6.907344e-01 & 1.049444e-28 & 1.519317e-28 \\ 
3 & 4 & 4 &  2.903904e-01 & 8.442770e-26 & 2.907386e-25 \\ 
3 & 4 & 5 &  5.333668e-01 & 5.987514e-25 & 1.122588e-24 \\ 
\hline
\end{tabular}
\end{center}
\label{fig::allRegSolvability}
\caption
    {
      We show the lowest value of
      $\cali$ for an entire Cartesian grid of size $32^D$
      with all
      boundary conditions homogeneous Neumann.
      We vary Taylor power $P^T$, weighting
      function exponents $P^W$, and dimensionality $D$. 
    }
\end{table}
\end{small}

@@

For context, allow us to recap a relevant bit of finite volume formalism.
Consider a volume $\ibold$ bounded by the set of faces
$\{f\}  = \partial \ibold$, where the outward-facing normal $\nhat$ of
each face is known, as is its  $V_\ibold$.
Given a flux function $F$, the average divergence $< \nabla \cdot F >$
is given by the divergence theorem:
\begin{equation}
\begin{array}{l l}
< \nabla \cdot F > &= \frac{1}{V_\ibold} \int_\ibold \nabla \cdot F dV \\
           &= \frac{1}{V_\ibold} \int_{\partial \ibold} F\cdot \nhat dA \\
           &= \frac{1}{V_\ibold} \sum_{f \in \partial
  \ibold}\int_{\partial \ibold} F\cdot \nhat dA.
\end{array}
\label{eqn::discreteDivergence}
\end{equation}
For all coordinate-aligned faces, the section above's description for how to
compute the flux is adequate.   What remains are fluxes through cut
faces.

@@

\subsection{The effect of Dirichlet boundary conditions on
  system matrix conditioning} In section \ref{sec::openface}, we
discuss how adding Neumann boundary condition equations can improve
the conditioning of the system matrix.  When the boundary conditions
are set to Dirichlet, we get similar results.  See tables
\ref{fig::mergedDiri}, \ref{fig::unmergedDiri}, and
\ref{fig::allRegDiri} for details.  Again, boundary conditions help
stabilize the system but it is still the case that aggressive choices
of $P_T$ or $P_W$ can result in system matrices that cannot be
credibly inverted without very high precision numerical arithmetic.
Again, this seems less a matter of a particular grid configuration and
more  a property of the mathematical framework used to generate the
stencil.

\begin{small}
\begin{table}
\begin{center}
\begin{tabular}{|ccc|ccc|} \hline
 $D$ & $P^T$  & $P^W$  & $\lambda_{max}$ & $\lambda_{min}$   & $\cali$ \\
 \hline
2 & 1 & 1 &  1.537125e+00 & 1.150772e-03 & 7.486517e-04 \\ 
2 & 1 & 2 &  4.940953e-01 & 9.196245e-05 & 1.861229e-04 \\ 
2 & 1 & 3 &  3.324177e-01 & 1.064946e-05 & 3.203639e-05 \\ 
2 & 1 & 4 &  2.853322e-01 & 1.705230e-06 & 5.976295e-06 \\ 
2 & 1 & 5 &  2.665812e-01 & 3.349335e-07 & 1.256403e-06 \\ 
2 & 2 & 1 &  2.407807e+00 & 1.457063e-06 & 6.051408e-07 \\ 
2 & 2 & 2 &  4.940956e-01 & 5.469314e-08 & 1.106934e-07 \\ 
2 & 2 & 3 &  3.324177e-01 & 2.879583e-09 & 8.662545e-09 \\ 
2 & 2 & 4 &  2.853323e-01 & 1.628939e-10 & 5.708921e-10 \\ 
2 & 2 & 5 &  2.665812e-01 & 1.084310e-11 & 4.067466e-11 \\ 
2 & 3 & 1 &  2.407807e+00 & 5.822902e-10 & 2.418342e-10 \\ 
2 & 3 & 2 &  9.272889e-01 & 4.520282e-11 & 4.874729e-11 \\ 
2 & 3 & 3 &  3.324177e-01 & 1.618058e-12 & 4.867544e-12 \\ 
2 & 3 & 4 &  2.853323e-01 & 7.227717e-14 & 2.533088e-13 \\ 
2 & 3 & 5 &  2.665812e-01 & 3.194386e-15 & 1.198279e-14 \\ 
2 & 4 & 1 &  2.943604e+00 & 6.445948e-26 & 2.189815e-26 \\ 
2 & 4 & 2 &  9.877639e-01 & 6.302117e-27 & 6.380185e-27 \\ 
2 & 4 & 3 &  6.561304e-01 & 6.778809e-28 & 1.033150e-27 \\ 
2 & 4 & 4 &  2.868930e-01 & 5.733222e-27 & 1.998383e-26 \\ 
2 & 4 & 5 &  5.327046e-01 & 5.117605e-25 & 9.606835e-25 \\
\hline
3 & 1 & 1 & 4.639099e+00 & 3.887142e-03 & 8.379088e-04 \\ 
3 & 1 & 2 & 1.079504e+00 & 2.369894e-04 & 2.195354e-04 \\ 
3 & 1 & 3 & 6.747118e-01 & 2.359590e-05 & 3.497182e-05 \\ 
3 & 1 & 4 & 5.721834e-01 & 3.540976e-06 & 6.188534e-06 \\ 
3 & 1 & 5 & 5.333668e-01 & 6.792558e-07 & 1.273525e-06 \\ 
3 & 2 & 1 & 4.639156e+00 & 3.176883e-06 & 6.847975e-07 \\ 
3 & 2 & 2 & 1.079505e+00 & 1.512048e-07 & 1.400686e-07 \\ 
3 & 2 & 3 & 6.747119e-01 & 6.991076e-09 & 1.036157e-08 \\ 
3 & 2 & 4 & 5.721835e-01 & 3.641527e-10 & 6.364265e-10 \\ 
3 & 2 & 5 & 2.669165e-01 & 1.146071e-11 & 4.293744e-11 \\ 
3 & 3 & 1 & 4.639157e+00 & 1.340233e-09 & 2.888959e-10 \\ 
3 & 3 & 2 & 1.079505e+00 & 7.151353e-11 & 6.624659e-11 \\ 
3 & 3 & 3 & 6.747119e-01 & 3.977531e-12 & 5.895155e-12 \\ 
3 & 3 & 4 & 2.869943e-01 & 5.757645e-14 & 2.006188e-13 \\ 
3 & 3 & 5 & 2.669165e-01 & 1.723124e-15 & 6.455665e-15 \\ 
3 & 4 & 1 & 3.106689e+00 & 3.897434e-26 & 1.254530e-26 \\ 
3 & 4 & 2 & 1.171168e+00 & 3.832896e-28 & 3.272712e-28 \\ 
3 & 4 & 3 & 7.002071e-01 & 3.537134e-28 & 5.051555e-28 \\ 
3 & 4 & 4 & 2.903904e-01 & 8.048694e-26 & 2.771680e-25 \\ 
\hline
\end{tabular}
\end{center}
\label{fig::allRegDiri}
\caption
    {
      We show the lowest value of
      $\cali$ for an entire Cartesian grid of size $32^D$ with all
      boundary conditions homogeneous Dirichlet.
      We vary Taylor power $P^T$, weighting
      function exponents $P^W$, and dimensionality $D$.
    }
\end{table}
\end{small}
